Characteristic polynomials of p-adic matrices

### 2017-02-04 charpoly/charpoly.tex
Contributions, notations

\subsection*{The contribution of this paper.} % Or \textbf

@@

\subsection*{Notation} Throughout the paper, $K$ will refer to a complete,
discrete valuation field, $\val : K \twoheadrightarrow \Z \cup \{+\infty\}$ to its valuation,
$\OK$ its ring of integers and $\pi$ a uniformizer. \todo{$\softO$, $\omega$, scalar matrices and $I_n$}

>>>>

\subsection*{The contribution of this paper.}

Thanks to the application the framework of differential precision of \cite{caruso-roe-vaccon:14a} in \cite{caruso-roe-vaccon:15a},
we known that
the precision on the characteristic polynomial
$\chi_M$ of a matrix $M \in M_n(\mathbb{Q}_p)$
is determined by the comatrix $com(X I_n -M).$
In this article, we provide:
\begin{enumerate}
\item A factorization of $com(X I_n -M)$ (when $M$ has a cyclic vector).
\item A simple criterion to decide whether there can be a gain in precision
on $\chi_M$ when $M \in M_n(\mathbb{Z}_p).$
\item A $\softO (n^3)$ algorithm with operations in $O_K$
to compute the optimal precision on each coefficient of $\chi_M$ when $M$
is given with uniform precision on its entries.
\item A $\softO (n^\omega)$ algorithm to compute the 
optimal precision on each coefficient of $\chi_M$
for $M \in M_n (K),$
using divisions.
\end{enumerate}

@@

\subsection*{Notation} Throughout the paper, $K$ will refer to a complete,
discrete valuation field, $\val : K \twoheadrightarrow \Z \cup \{+\infty\}$ to its valuation,
$\OK$ its ring of integers and $\pi$ a uniformizer. 
We will write that $f(n)=\softO (g(n))$ if there exists some
$k \in \mathbb{N}$ such that $f(n)=O(g(n) \log (n)^k.$
We fix an $\omega \in \mathbb{R}$ such that the multiplication
of two matrices over a ring is in $O(n^\omega)$ operations in the
ring. Currently, the smallest proved $\omega$ is less than
$2.3728639 $ thanks to \cite{legall:14a}. 
$I_n$ will denote the identity matrix of rank $n$ in $M_n(K).$
When there is no ambiguity, we will drop this $I_n$
for scalar matrices, \textit{e.g.} for $\lambda \in K$
and $M \in M_n (K),$ $\lambda-M$ denotes $\lambda I_n -M.$
